% !TeX root = ../main.tex

\sysusetup{
  keywords  = {Agent Skill；条件排序；SkillRank；三维语义地形；实验复现；系统原型},
  keywords* = {Agent Skill; conditional ranking; SkillRank; 3D semantic terrain; experiment reproducibility; system prototype},
}

\begin{abstract}
  随着大语言模型驱动的智能体（Agent）不断走向真实软件开发、文档处理与自动化运维场景，围绕 Agent Skill 的公共生态也在快速膨胀。Skill 本质上是一个以 Description 为定位线索、以 Content 为实际注入内容的能力单元；当候选 Skill 从几十条增长到数千条后，用户和 Agent 都难以稳定判断“当前任务真正应该使用哪几条 Skill”。现有公开目录普遍依赖 GitHub Star、下载量或人工分类进行全局排序，但这些信号既无法刻画 Skill 所处的任务语义邻域，也无法有效反映 Content 本身的结构质量。

  针对上述问题，本文设计并实现了 Taste 系统原型及其核心排序算法 SkillRank。本文将 Skill 建模为二元组 $\langle Description, Content \rangle$，提出 Skill 价值评估应被视为一个条件排序问题：只有在 Description 语义相近的竞争邻域内部，对 Content 质量进行排序才有意义。基于这一观点，SkillRank 先在 Description 空间中完成语义聚类，再在各聚类内部以 Content embedding 构建稀疏 K-NN 亲缘图，通过 Content 原型中心度与 PageRank 风格的图传播分融合得到最终质量分；同时使用 FAISS 近似最近邻搜索为更大规模数据保留可扩展实现路径。

  在实验层面，本文整理了 10028 条 GitHub Skill 样本（覆盖 2430 个仓库）作为统一数据资产，并按当晚实验预算仅运行 100 和 1000 两个规模的子集。评测采用聚类内 GitHub 仓库 Star 排序作为弱监督 proxy，以加权 within-cluster Spearman $\rho$ 衡量排序质量。经 AutoResearch 风格单变量迭代调优后，SkillRank 在 100 规模上由初始配置的 $-0.0084$ 提升至 $0.0600$，在 1000 规模上由 $0.0464$ 提升至 $0.0562$；两组结果均优于最强非图结构基线 Content centroid（分别为 $0.0284$ 与 $0.0527$）。效率对比表明，在 100 和 1000 这两个小中规模下，精确 K-NN 的常数开销仍低于 FAISS，但近似检索造成的质量损失接近于零。

  在系统实现层面，本文完成了从离线实验流水线到在线演示 Demo 的完整闭环：离线部分使用 Python、sentence-transformers、scikit-learn、SciPy 与 FAISS 生成聚类、SkillRank 分值和三维语义地形产物；在线部分以 Plotly.js \texttt{scatter3d} 提供真实三维地形展示，并通过 TF-IDF 候选召回结合预计算 SkillRank 分值完成查询重排。全部实验日志、汇总 JSON、截图与 Demo 产物均被保留，以支持后续迭代与复现。
\end{abstract}

\begin{abstract*}
  As large language model driven agents move into real software engineering, documentation, and automation workflows, the public ecosystem of Agent Skills is expanding rapidly. A Skill can be viewed as a capability unit with Description as the routing cue and Content as the actual injected instruction. Once the number of available Skills grows from dozens to thousands, both users and agents struggle to decide which Skills should be activated for a given task. Existing public directories mainly rely on GitHub Stars, download counts, or manual categories for global ranking, but these signals fail to capture semantic competition neighborhoods and do not directly measure the structural quality of Skill Content.

  To address this problem, this thesis designs and implements the Taste system prototype together with its core ranking algorithm, SkillRank. Skills are formalized as tuples $\langle Description, Content \rangle$, and Skill value evaluation is framed as a conditional ranking problem: Content quality should only be ranked within semantically comparable neighborhoods defined by Description. Based on this formulation, SkillRank first clusters Skills in the Description space, then builds sparse K-NN affinity graphs within each cluster using Content embeddings, and finally fuses a content centroid prior with a PageRank-style graph propagation score. FAISS approximate nearest neighbor search is incorporated to preserve a scalable implementation path for larger datasets.

  For evaluation, this thesis consolidates a unified asset of 10,028 GitHub Skill samples from 2,430 repositories, and, under the execution budget for the current experiment session, runs two scales only: 100 and 1000. Weighted within-cluster Spearman $\rho$ is used as the ranking metric, with repository GitHub Stars serving as a weak supervision proxy inside each cluster. After AutoResearch-style single-parameter iterative tuning, SkillRank improves from $-0.0084$ to $0.0600$ on the 100-sample setting and from $0.0464$ to $0.0562$ on the 1000-sample setting. Both tuned results outperform the strongest non-graph baseline, namely the Content centroid prior ($0.0284$ and $0.0527$ respectively). Runtime comparisons further show that, at these small and medium scales, exact K-NN is still faster than FAISS due to constant-factor overhead, while the ranking-quality gap remains negligible.

  At the system level, the thesis delivers an end-to-end loop from offline experimentation to an online interactive demo. The offline pipeline uses Python, sentence-transformers, scikit-learn, SciPy, and FAISS to generate clusters, SkillRank scores, and semantic terrain artifacts. The online demo renders a true 3D semantic terrain with Plotly.js \texttt{scatter3d}, and uses TF-IDF candidate recall followed by SkillRank-based reranking for query interaction. All experiment logs, summary JSON files, screenshots, and demo artifacts are preserved to support future iterations and reproducibility.
\end{abstract*}
